---
title: \vspace{-1cm}"Practical Machine Learning Project - Prediction"
author: "Ulises Quiñonez"
date: "August 2020"
output:
  pdf_document: default
  html_document:
  keep_md: yes
subparagraph: yes
header-includes: |
  \usepackage{titlesec}
  \titlespacing{\section}{0pt}{10pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
  \titlespacing{\subsection}{0pt}{10pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
  \titlespacing{\subsubsection}{0pt}{10pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
---
# Excecutive Summary
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

The goal of this project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. At the end of this project the prediction model is used to predict 20 different test cases.

```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
options(warn=-1)
fig.path = "instructions_fig/"
```

# Loading the packages to use
```{r results='hide', message=FALSE, warning=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```

# Downloading the data
```{r}
trUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
tsUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trFile <- "training.csv"
tsFile  <- "testing.csv"
if (!file.exists(trFile)) {
  download.file(trUrl, destfile=trFile, method="curl")
}
if (!file.exists(tsFile)) {
  download.file(tsUrl, destfile=tsFile, method="curl")
}
```  

# Reading the Data
```{r, cache = T}
trainRaw <- read.csv("training.csv")
testRaw <- read.csv("testing.csv")
dim(trainRaw)
dim(testRaw)
```
Training dataset:
19622 observations
160 variables
Testing data set contains
20 observations 
160 variables 
Variable to predict: classe 

# Cleaning the data
```{r, cache = T}
sum(complete.cases(trainRaw))
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
```
Training dataset:
19622 observations
53 variables
Testing data set contains
20 observations 
53 variables 

# Spliting the data
```{r, cache = T}
set.seed(123) 
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```
 
# Modeling the data
We use the Random Forest algorithm with 5-fold cross validation and figure out the performance of the model:
```{r, cache = T}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf
predictRf <- predict(modelRf, testData)
confusionMatrix(testData$classe, predictRf)
accuracy <- postResample(predictRf, testData$classe)
accuracy
oose <- 1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
oose
```
It's +99% with an estimated error of 0.58%.

# Predicting
```{r, cache = T}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result
```  

# And graphing:

Correlation:
```{r, cache = T}
corrPlot <- cor(trainData[, -length(names(trainData))])
corrplot(corrPlot, method="color")
```

Decision Tree:
```{r, cache = T}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel)
```
